---
title: "Lab 01"
author: "Anonymous"
date: "2023-08-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Text Preliminaries 
```{r text_preliminaries}
little_women <- readLines("./data/little_women.txt")
lines <- length(little_women)
lw_clean <- little_women[little_women !=""]
lw_text <- paste(lw_clean, collapse = " ")
```
## Lab Answers
1. There are 47 chapters and 21,022 lines in Little Women. 
2. Done in the above. 
3. Chapter splitting done below:
```{r text_splitting}
lw_chaps <- strsplit(lw_text, "CHAPTER ")
lw_chap <- lw_chaps[[1]] #Why does this thing count the blurb at the start with the Project Gutenburg 
#as a chapter? 
lw_chapters <- lw_chap[2:length(lw_chap)]

check_length <- length(lw_chapters)
```
Note that the length of `lw_chapters` reflects the correct number of chapters. The only thing I don't understand is why I have to remove the first element from my list, as the phrase "CHAPTER " does not appear in the Project Gutenburg blurb, but seems to have been included in `lw_chap`. 

### Counting and Storing Appearances 
```{r counting}
library(stringr) #loading the stringr library
#counting appearances in each chapter
count_Amy <- str_count(lw_chapters, "Amy")
count_Beth <- str_count(lw_chapters, "Beth")
count_Jo <- str_count(lw_chapters, "Jo")
count_Laurie<- str_count(lw_chapters, "Laurie")
count_Meg <- str_count(lw_chapters, "Meg")
  
#placing cumulative counts in a data frame for easier plotting
counting <- data.frame(
  cumul_count_Amy = cumsum(count_Amy), 
  cumul_count_Beth = cumsum(count_Beth), 
  cumul_count_Jo = cumsum(count_Jo), 
  cumul_count_Laurie = cumsum(count_Laurie), 
  cumul_count_Meg = cumsum(count_Meg)
)
```

### Required Plot
```{r plot_cumulative, echo=FALSE}

colours <- c("indianred3", "darkgoldenrod2", "cornflowerblue", "darkolivegreen3", "mediumpurple3")
matplot(counting, type = "l",
        main = "cumulative counts of character appearances",
        xlab = "chapters",
        ylab = "cumulative appearances",
        col = colours,
        lwd = 2, 
        lty = 1)

legend('topleft', 
       legend = c("Amy", "Beth", "Jo", "Laurie", "Meg"), 
       col = colours, 
       lty = 1, 
       title = 'characters',
       cex =.8)
```

