---
title: "lab_02"
output: pdf_document
date: "2023-09-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries
Here, we are reading the CSV:
```{r cal_football_season}
cal_fb <- read.csv("./data/cal_fb.csv")

```

Here, we remove the first column of the dataframe above. Then, we write a function that calculates Cal's total score and the opponent's total score from a single row of scores. 
```{r quarter_scores}
quarter_scores <- cal_fb[, -1]
sum_scores <- function(x) {
  row_scores <- x
  cal_scores <- sum(row_scores[1:4])
  opponent_scores <- sum(x) - cal_scores
  return(c(cal_scores, opponent_scores))
}
```
We check if our function returns the correct values for the first row: 
```{r checking_sum, echo = FALSE}
sum_scores(quarter_scores[1,])
```

Now, we want to create a new dataframe that shows the final scores against all the opponents. 

```{r final_scores}
first_column <- cal_fb[ ,1]
#apply presents the results as two rows, so we have to flip the rows and columns
second_and_third_columns <- t(apply(quarter_scores, 1, sum_scores))

#storing the final scores as a data frame
final_scores <-as.data.frame(cbind(first_column, second_and_third_columns))

```

I used `dplyr` to change the column names, but have hidden the code for the PDF file. Please look at the `markdown` file for more details. 
```{r changing_col_names, echo = FALSE}
#changing the column names to the appropriate ones
suppressPackageStartupMessages(library("dplyr"))
final_scores <- final_scores %>%
              rename("Cal_Score" = "V2",
                      "Opponent_Score" = "V3")

```

## Is Cal a Loser 
Here, we write a function that checks if Cal football loses against its opponent. We don't account for draws in this function, but I guess it's fine since none of the scores are a draw anyway. 
```{r cal_lost}
did_cal_lose <- function(x) {
  #Not sure why I had to put it in an if-else statement
  #Previously wrote as return(x[2] < x[3]), but then it would also print 
  #Cal_score row number for some reason
  #have to change the results from string to integer, or it will compare index by index
  if (strtoi(x[2]) < strtoi(x[3])) {
    return(TRUE)
  }
  else {
    return(FALSE)
  }
}

```

Checking to see if it returns the right result, for row 6:
```{r check_lose, echo = FALSE}
did_cal_lose(final_scores[6,])
```

## Carlos is Nosy

Now, we log the results that Carlos is looking for: 
```{r results_tabulation}
results <- apply(final_scores, 1, did_cal_lose)
final_scores_with_results <- cbind(final_scores, results)

cal_wins <- sum(final_scores_with_results[,4] == FALSE)
cal_losses <- sum(final_scores_with_results[,4] ==TRUE)
```

Let's check if the wins and losses are correct, where the first number is the number of wins and the second one is the number of losses:
```{r show_results, echo = FALSE}
print(cal_wins)
print(cal_losses)
```

## Head or Tails (Sonic likes Tails)
Here, we write a function that tosses a coin once as well another function that tosses the coin some set amount that we want, and then stores the results.
```{r coin_toss}
#tosses the coin once
toss <- function(x) {
  sample(c("H", "T"), size = 1)
}

#tosses the coin multiple times, according to the argument supplied
toss_multiple <- function(x) {
  store <- numeric(x)
  for (i in 1:x) {
    store[i] <- toss()
  }
  return(store)
}

```


Here, we have written a function that counts the number of heads in a given number of coin tosses. 
```{r head_count}
count_heads <- function(x){
  sum(grepl("H", toss_multiple(x)))
}

```

Now, we toss the coin 100 times in 1000 simulations by using the `sapply`, and obtain the required histogram. 
```{r actual_head_count, echo = FALSE}
wrapper_heads<- function(x){
 count_heads(100)
}

number_of_heads <- sapply(1:1000, wrapper_heads)

```

```{r head_plot, echo = FALSE}
hist(number_of_heads,
     main = "Number of Heads Appearing in 100 Tosses in 1000 Simulations",
     xlab = "Number of Heads in 100 Tosses",
     col = "cornflowerblue",
     ylim = c(0,200),
     breaks = 20)
```

We see that the most frequent number of heads appearing is around 50 or so, which makes sense, since there's a 50% chance of heads appearing. 
