---
title: "lab_04"
output: pdf_document
date: "2023-09-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Vote Now for Your Favourite Thai Restaurant

Here, we are reading the `.csv` file and determining the sample size we need:
```{r preliminaries}
real_votes <- read.csv("./data/votes.csv")
sample_size <- nrow(real_votes)

```


Now, we write the required funciton that simulates one bootstrap sample and returns Imm Thai's percentage of votes:
```{r bootstrapping_thai_once}
one_resampled_percentage <- function(){
  bootstrap <- real_votes[sample(sample_size, replace = TRUE),]
  return((sum(bootstrap == "Imm Thai")/sample_size)*100)

}
```

Now, we write the required function:
```{r multiple_bootstrap}
wrapper_percentage <- function(x){
  one_resampled_percentage()
}
percentages_in_resamples <- function(m){
  return(sapply(1:m, wrapper_percentage))
}
```

Here, we simulate 2000 times as required, and find the lower and upper bounds for the 95%-confidence interval:
```{r 2000_sim}
so_many_sim <- percentages_in_resamples(2000)
imm_lower_bound <- quantile(so_many_sim, 0.025)
imm_upper_bound <- quantile(so_many_sim, 0.975)
```

Our histogram for the simulations:

```{r hist_sim, echo = FALSE}
hist(so_many_sim,
     main = "Histogram of Imm Thai's Percentage of Votes",
     xlab = "Imm Thai's Percentage of Votes",
     col = "hotpink",
     breaks = 15)
abline(v = c(imm_lower_bound, imm_upper_bound), lwd = 3, col = "darkgoldenrod1")

```

## Imm (Thai) Doing So Much Better Than You

Now, we write the function for calculating Imm Thai's lead over the other restaurants: 
```{r difference}
one_resampled_difference <- function() {
  bootstrap <- real_votes[sample(sample_size, replace = TRUE),]
  imm_thai_percentage <- (sum(bootstrap == "Imm Thai")/sample_size)*100
  lucky_house_percentage <- (sum(bootstrap == "Lucky House")/sample_size)*100
  thai_temple_percentage <- (sum(bootstrap == "Thai Temple")/sample_size)*100
  thai_basil_percentage <- (sum(bootstrap == "Thai Basil")/sample_size)*100
  return(imm_thai_percentage - (lucky_house_percentage
                                + thai_temple_percentage 
                                + thai_basil_percentage))
}
```

Simulating 2000 times:
```{r multi_difference}
wrapper_difference <- function(x){
  one_resampled_difference()
}

resampled_leads <- sapply(1:2000, wrapper_difference)
```

Finally, the required histogram, in a lovely blue:

```{r hist_difference, echo = FALSE}
hist(resampled_leads, 
     main = "Histogram of 2000 Bootstrap Samples of Imm Thai's Lead",
     xlab = "Leads",
     col = "deepskyblue4",
     breaks = 15,
     xlim = c(-5,15))

```

And here is question 7: 
```{r lead_ci, echo = FALSE}
diff_lower <- quantile(resampled_leads, 0.025)
diff_upper <- quantile(resampled_leads, 0.975)

print(noquote(paste0(
  "A 95%-CI for Imm Thai’s true lead over the other restaurants is (", round(diff_lower, digits = 2), ", ", round(diff_upper, digits = 2), ")."
)))
```

